Matching ontologies with distributed word embeddings

### 2016-07-07 Paper/Draft.tex
Added images to algorithms

@@

 \makebox[\textwidth][c]{
 \includegraphics[width=1.35\textwidth]{plots/miceman}
 }\footnote{Credit: Sanna Dinh}
 \newpage

 \pagenumbering{roman}
 \section*{Acknowledgements}
 Arjen de Vries, professor at the Radboud University of Nijmegen and my mentor - Kennis enthusiasme en als het tegenzat kon hij me toch motiveren.
 Sanna Dinh, my partner - Nachtenlang opgebleven om het schrijfwerk en plaatjes
 Jeroen Vuurens, Delft University of Technology - Sparren over word2vec
 Xander Wilcke - Inspireren over het Semantisch Web
 Marianne - op de gekste tijdstippen skypen en proof-readen
 Suzan Verberne, professor at the Radboud University of Nijmegen - Nakijken
 Flynn, my pug - For the fresh air when I needed a minute to relax

 \section*{Abstract}
 McCallum proposed using distributed word embeddings for knowledge base completion, the problem of finding relations \textit{within} a knowledge base. In this study distributed word embeddings were used for ontology matching, the problem of finding relations \textit{between} knowledge bases. To test if distributed word embeddings also work on ontology matching, multiple algorithms that adapt these embeddings to ontology graphs were implemented.

 The conclusion is that distributed word embeddings can be applied to ontology matching. The performance is not as good as the state-of-the-art methods yet. However, distributed word embeddings have potential to improve with more data as opposed to current methods which are hand-crafted.

 \section*{Summary}
 \tableofcontents
 \newpage

 %Vragen
 % Hypothese waarom ik dat denk? Antwoord see Arjen notes
 % extra hypothese. Antwoord see Arjen notes (I think? otherwise its: don't add ensemble hypothesis)

@@

 An example: since the Hummingbird update, when someone Googles "Who is Barack Obama" they will find a part of the Knowledge Graph, see Figure \ref{hummingbird}.. This system improves web searching by finding answers rather than pages. 
 \paragraph{}
 The many uses of ontology matching and its role in the Semantic Web show that it is important to research all avenues that might improve ontology matching, as it will have a direct result in many fields.

@@

 \subsection{Algorithms} \label{implementations}
 In this section, I will discuss the different implementations that were made of the algorithm. All implementations use the same basic algorithm in different ways by taking into account the context in different ways. All of these implementations were tested and the results can be found in Section \ref{results}. Some implementations extend others, but all implementations are covered to ensure completeness. This way, all improvements are recorded precisely.
 All algorithms use word embeddings, though the way they use word embeddings is different.  The explanations refer back to the example in Section \ref{example}.
  \subsubsection{Pre-trained models}
  The simplest way to use node embeddings for ontology matching is to obtain a model already trained on normal text. This model should already have learned some semantic context, although neighbourhood in text is different from that in an ontology. Also, since the words it is trained on might not match the labels from the ontology, two things need to be done. Firstly, if a word is not in the model, it is ignored. Secondly, the labels are split up into separate words and the resulting vectors are averaged.
  This model was chosen because of its ease of implementation and to see if the graph-based context improves ontology matching over the text-based context.
  \subsubsection{DeepWalk}
  DeepWalk generates sentences by randomly walking through the ontology and processes those sentences using a normal word embedding model. This means it is possible to have a wide neighbourhood, as the neighbours of neighbours are also used as context, or even further neighbours depending on the scope of the algorithm. It also allows for weighting of edges to change the sampling rate of the random walk, which can be used to prioritize important nodes or highlight informative neighbours.
  One adaptation over normal Deepwalk that was considered is switching between ontologies when on a concept from the correspondence set. For example, if both ontologies contain \textit{eye}, when the algorithm lands on the corresponding node, it will have a chance to switch to the other ontology. As such the sentence will contain concepts from both ontologies and they will move closer together. See \ref{sentences} for an example of the sentences generated by DeepWalk.
  \subsubsection{Node vectorisation}
  The most basic \emph{Node2vec} model converts a node to a vector purely based on its own label. This method has one advantage over simple string matching: if no node is found in the other ontology that matches exactly, we can still find a node that is similar since it is close in the vector space. Therefore, this method should already be an improvement over the most basic string matching algorithms. It may even be competitive with more advanced string matchers (that look at substrings of labels). For example, \emph{whisker} and \emph{moustache} may be matched by a substring matcher, but are also semantically similar since they will often be mentioned in the same context.
  This method was chosen because it is the foundation of all other models that use the graph-context. It is required to see if the extensions improve results.
  \subsubsection{Bidirectional connections}
  More connections means bigger context. For neural networks more data is always a good things. Therefore it is likely that this will improve results somewhat.
  \subsubsection{Select from senses}
  One way to take into account the influence of neighbouring nodes is to train a multi-sense node embedding model as described in Section \ref{multisense}, and selecting from the different node senses based on which is closest to the context average.
  \subsubsection{Hot start}
  It is hard to classify the proposed algorithm in terms of the conventional ontology alignment method classes. It is not purely string matching, since it takes into account information from the context. However, it is also not structural, as it can work without a seed alignment and uses information other than the structure as well. It is not semantic (or logical) since it does not use inference. Nor is it terminological, which looks at dictionaries or other ontologies to find matches. However, it can use dictionaries and other ontologies to improve performance.
  The algorithm can find matches based purely on the given node and edge labels. However the algorithm might benefit from a hot start. Such a seed alignment might allow the algorithm to train with certain words that are known synonyms as if they were the same word, thus increasing the number of training samples per word (on known relevant words) and decreasing sparsity. This may help the model, improving performance. However, since some information from other algorithms is now used, it would not be fair to compare the results to the cold start algorithm results. For example, it may just add the results from the generic string matcher to its own results, improving performance, without learning any new relationships. A new testing method had to be designed for this algorithm. This method would have to compare it to the string matchers, to see if it improved over their results, rather than compare it to the cold start algorithm. However, if we compare the cold and hot algorithms with the generic string matchers, we may be able to compare them indirectly.
  An example of a hot start is that \emph{mouse torso} is already matched with \emph{human torso}. Then it is likely that their neighbours are also related. Since the \emph{front paws} and \emph{arms} are next to the \emph{torso}, they are more likely to be matched now. In the case of \emph{Node2vec} they are more likely to be matched, since their neighbourhoods are the same (the only neighbour of both \emph{arms} and \emph{paws} is \emph{connectedTo torso}).

@@

  \subsection{F-score}
  \begin{table}[H]
  \caption[$F_1$]{The $F_1$-scores for the different matchers for the mice and medical datasets compared to the result of AML at OAEI 2015 \cite{oaei2015}}.
  \label{fscores}
  \begin{tabular}{l|ll}
  \multirow{2}{*}{\textbf{Matcher}} & \multicolumn{2}{c}{\textbf{$F_1$-score}} \\
  & Anatomy & Medical\\
  \hline
  Label Match & 0.6926 & 0.4905 \\
  Random & 0.6926 & 0.4905 \\
  Pre-trained & 0.6926 & 0.3874 \\
  Bidirectional & 0.8432 & 0.4946 \\
  Multi-sense & 0.6926 & 0.4905 \\
  Deepwalk & 0.6926 & 0.4906 \\
  Alternative DeepWalk & 0.6926 & 0.4905 \\
  \textbf{AML ensemble} & \textbf{0.94} & \textbf{0.81} \\
  \end{tabular}
  \end{table}

 \newpage
 \subsection{Interpretation}

@@

 To answer this question I have examined the following points from Section \ref{hypotheses} for every algorithm (Pre-trained Word2vec, DeepWalk, Node2vec):

@@

 \newpage
 \bibliographystyle{plain}
 \bibliography{bibliography}{}

>>>>

 \hfuzz=20pt % No more overfull hboxes!

@@

 \makebox[\textwidth][c]{
 \includegraphics[width=1.35\textwidth]{plots/miceman}
 }\footnote{Credit: Sanna Dinh}

 \newpage

 \pagenumbering{roman}
 \section*{Acknowledgements}
 I would like to thank the following people:
 \paragraph{}
 \paragraph{}
 Arjen de Vries, professor at the Radboud University of Nijmegen and my mentor, for his limitless pool of knowledge and enthusiasm. For getting me out of the mud when I was stuck and motivating me to push on. 
 \paragraph{}
 Suzan Verberne, professor at the Radboud University of Nijmegen and co-reader, for helping me start and finish this thesis.
 \paragraph{}
 Jeroen Vuurens, from Delft University of Technology, for helping me figure out Word2vec.
 \paragraph{}
 Xander Wilcke, from the VU Amsterdam, for inspiring me to combine Machine Learning with the Semantic Web.
 \paragraph{}
 Sanna Dinh, my partner, for spending countless late hours making this thesis readable and designing many of the images, making this thesis understandable.
 \paragraph{}
 Marianne, my mother-in-law, for being available on Skype at the most unconventional times to advise and proof-read. 
  \paragraph{}
 Flynn, my pug, for the fresh air during our walks when I needed a minute to relax.

 \newpage 
 \begin{abstract}
 McCallum proposed using distributed word embeddings for knowledge base completion, the problem of finding relations \textit{within} a knowledge base. In this study distributed word embeddings were used for ontology matching, the problem of finding relations \textit{between} knowledge bases. To test if distributed word embeddings also work on ontology matching, multiple algorithms that adapt these embeddings to ontology graphs were implemented.

 The conclusion is that distributed word embeddings can be applied to ontology matching. The performance is not as good as the state-of-the-art methods yet. However, distributed word embeddings have potential to improve with more data as opposed to current methods which are hand-crafted.
 \end{abstract}

 \newpage
 \tableofcontents
 \newpage

@@

 An example: since the Hummingbird update, when someone Googles "How old is Barack Obama" they will find a part of the Knowledge Graph, see Figure \ref{hummingbird}. This system improves web searching by finding answers rather than pages.

 \begin{figure}
 \includegraphics[width=\textwidth]{plots/hummingbird}
 \caption[Hummingbird search]{A google search after the Hummingbird update returns a triple from Google's knowledge base, the Knowledge Graph.}
 \label{hummingbird}
 \end{figure}

 The many uses of ontology matching and its role in the Semantic Web show that it is important to research all avenues that might improve ontology matching, as it will have a direct result in many fields.

@@

 \subsection{Algorithms} \label{implementations}
 In this section, I will discuss the different implementations that were made of the algorithm. All implementations use the same basic algorithm in different ways by taking into account the context in different ways. All of these implementations were tested and the results can be found in Section \ref{results}. Some implementations extend others, but all implementations are covered to ensure completeness. This way, all improvements are recorded precisely.
 All algorithms use word embeddings, though the way they use word embeddings is different.  The explanations refer back to the example in Section \ref{example}.
  \subsubsection{Pre-trained models}
  The simplest way to use node embeddings for ontology matching is to obtain a model already trained on normal text. This model should already have learned some semantic context, although neighbourhood in text is different from that in an ontology. Also, since the words it is trained on might not match the labels from the ontology, two things need to be done. Firstly, if a word is not in the model, it is ignored. Secondly, the labels are split up into separate words and the resulting vectors are averaged.
  This model was chosen because of its ease of implementation and to see if the graph-based context improves ontology matching over the text-based context.
  \subsubsection{DeepWalk}
  DeepWalk generates sentences by randomly walking through the ontology and processes those sentences using a normal word embedding model. This means it is possible to have a wide neighbourhood, as the neighbours of neighbours are also used as context, or even further neighbours depending on the scope of the algorithm. It also allows for weighting of edges to change the sampling rate of the random walk, which can be used to prioritize important nodes or highlight informative neighbours.
  One adaptation over normal Deepwalk that was considered is switching between ontologies when on a concept from the correspondence set. For example, if both ontologies contain \textit{eye}, when the algorithm lands on the corresponding node, it will have a chance to switch to the other ontology. As such the sentence will contain concepts from both ontologies and they will move closer together. See Section \ref{sentences} for an example of the sentences generated by DeepWalk. See Figure \ref{embedding_alignment} for an example of two ontology graphs being mapped to the same space.

  \begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{plots/embedding_alignment}
  \caption[Two graphs embedded]{Two graphs with a few corresponding nodes are embedded into vector space. Because of the known correspondences the latent correspondences are also mapped close to each other.}
  \label{embedding_alignment}
  \end{figure}

  \subsubsection{Node vectorisation}
  The most basic \emph{Node2vec} model converts a node to a vector purely based on its own label. This method has one advantage over simple string matching: if no node is found in the other ontology that matches exactly, we can still find a node that is similar since it is close in the vector space. Therefore, this method should already be an improvement over the most basic string matching algorithms. It may even be competitive with more advanced string matchers (that look at substrings of labels). See Figure \ref{ontology_matrix}.

  \begin{figure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{plots/ontology_matching_matrix}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/ontology_matching_matrix2}
  \end{subfigure}
  \caption[Tensor with multiple ontologies]{The ontologies represented as a tensor. The connections (in green) are only within the ontologies, not between. The embeddings are used to predict new connections between the ontologies (in light green)}
  \label{ontology_matrix}
  \end{figure}

  An example, \emph{whisker} and \emph{moustache} may be matched by a substring matcher, but are also semantically similar since they will often be mentioned in the same context.
  This method was chosen because it is the foundation of all other models that use the graph-context. It is required to see if the extensions improve results.
  \subsubsection{Bidirectional connections}
  More connections means bigger context. For neural networks more data is always a good things. Therefore it is likely that this will improve results somewhat.
  \subsubsection{Select from senses}
  One way to take into account the influence of neighbouring nodes is to train a multi-sense node embedding model as described in Section \ref{multisense}, and selecting from the different node senses based on which is closest to the context average.
  \subsubsection{Hot start}
  It is hard to classify the proposed algorithm in terms of the conventional ontology alignment method classes. It is not purely string matching, since it takes into account information from the context. However, it is also not structural, as it can work without a seed alignment and uses information other than the structure as well. It is not semantic (or logical) since it does not use inference. Nor is it terminological, which looks at dictionaries or other ontologies to find matches. However, it can use dictionaries and other ontologies to improve performance.
  The algorithm can find matches based purely on the given node and edge labels. However the algorithm might benefit from a hot start. Such a seed alignment might allow the algorithm to train with certain words that are known synonyms as if they were the same word, thus increasing the number of training samples per word (on known relevant words) and decreasing sparsity. This may help the model, improving performance. However, since some information from other algorithms is now used, it would not be fair to compare the results to the cold start algorithm results. For example, it may just add the results from the generic string matcher to its own results, improving performance, without learning any new relationships. A new testing method had to be designed for this algorithm. This method would have to compare it to the string matchers, to see if it improved over their results, rather than compare it to the cold start algorithm. However, if we compare the cold and hot algorithms with the generic string matchers, we may be able to compare them indirectly.
  An example of a hot start is that \emph{mouse torso} is already matched with \emph{human torso}. Then it is likely that their neighbours are also related. Since the \emph{front paws} and \emph{arms} are next to the \emph{torso}, they are more likely to be matched now. In the case of \emph{Node2vec} they are more likely to be matched, since their neighbourhoods are the same (the only neighbour of both \emph{arms} and \emph{paws} is \emph{connectedTo torso}).

@@

  \subsection{F-score}
  \begin{table}[H]
  \caption[$F_1$]{The $F_1$-scores for the different matchers for the mice and medical datasets compared to the result of AML at OAEI 2015 \cite{oaei2015}}.
  \label{fscores}
  \begin{tabular}{l|ll}
  \multirow{2}{*}{\textbf{Matcher}} & \multicolumn{2}{c}{\textbf{$F_1$-score}} \\
  & Anatomy & Medical\\
  \hline
  Label Match & 0.6926 & 0.4905 \\
  Random & 0.6926 & 0.4905 \\
  Pre-trained & 0.6926 & 0.4905 \\
  Bidirectional & 0.8432 & 0.4946 \\
  Multi-sense & 0.6926 & 0.4905 \\
  Deepwalk & 0.6926 & 0.4906 \\
  Alternative DeepWalk & 0.6926 & 0.4905 \\
  \textbf{AML ensemble} & \textbf{0.94} & \textbf{0.81} \\
  \end{tabular}
  \end{table}

@@

 To answer this question I have examined the following points from Section \ref{researchquestion} for every algorithm (Pre-trained Word2vec, DeepWalk, Node2vec):

@@

 \newpage
 \bibliographystyle{unsrt}
 \bibliography{bibliography}{}
